---
title: "Assignment 3"
output: github_document
---

```{r}
library(gapminder)
library(tidyverse)
```

# Tasks 

## 1. Get the maximum and minimum of GDP per capita for all continents.
```{r}
minGDP <-gapminder %>% 
  group_by(continent,country) %>% 
  summarize(min=min(gdpPercap)) %>% 
  summarize(min=min(min)) 
knitr::kable(minGDP)
 
maxGDP <-gapminder %>% 
  group_by(continent,country) %>% 
  summarize(max=max(gdpPercap)) %>% 
  summarize(max=max(max)) 
knitr::kable(maxGDP)

  ## plot the minimum of GDP per capital for all continents
  ggplot(minGDP,aes(continent,min))+
  geom_point(aes(size=min,color=continent))
  
  
  ## plot the maximum of GDP per capital for all continents
    ggplot(maxGDP,aes(continent,max))+
  geom_point(aes(size=max,color=continent))
  
  
```

## 2. Look at the spread of GDP per capita within the continents

```{r}
ggplot(gapminder, aes(gdpPercap))+
  facet_wrap(~continent, scales="free_x","free_y")+
  geom_histogram(aes(fill=continent))
  
## we can also use combination of scatter plot and box plot to show the spread of the GDP per capita
ggplot(gapminder,aes(gdpPercap,continent))+
  geom_point(aes(colour=continent,size=gdpPercap),alpha=0.4)
```


## 3. Compute a trimmed mean of life expectancy for different years. Or a weighted mean, weighting by population.

the following is computed to find the 25% trimmed mean which is known as interquartile mean. The interquartile mean as well as the normal mean  of lifeExp for each country are found:

```{r}
temp <-gapminder %>% 
  group_by(year) %>% 
  summarize(trimmedMean=mean(lifeExp, trim=0.25),Mean=mean(lifeExp))
knitr::kable(temp)
```
# these results can be compared using the below plot:
```{r}
ggplot(temp)+
  geom_line(aes(year,y=trimmedMean),color="red")+
  geom_line(aes(year,y=Mean),color="blue")+
  xlab("year")+
  ylab("mean")
```


Compute a weighted mean of lifeExp for different years, weighting by population

```{r}
 gapminder%>% 
  group_by(year) %>% 
 summarize(weighted.mean(lifeExp,pop)) 

```


